---
title: "physiology_stats"
author: "Debbie Leung"
date: "2/11/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up
```{r}
#install.packages("Hmisc")
library("Hmisc")
#install.packages("corrplot")
#install.packages("olsrr")
#install.packages("car")
library(corrplot)
library(dplyr)
library(tidyr)
library(olsrr)
library(car) 
```

```{r}
my_data <- read.csv("physiology_cleaned.csv", stringsAsFactors = FALSE) %>% na_if("")
```

Steps:
1. Compute statistical tests
2. Compute normality and homogeneity tests
3. Compute any post-hoc tests for significant results

## Covariables with Covariables

### Correlation Matrix

Problem: 
cannot lump all continuous variables to display into a correlation matrix because of missing NA values... The matrices computed below removed avg_temp and temp_sd.
```{r}
con_var_na <- my_data %>% select(shell_length:heart_rate)
test <- cor(con_var_na, use = "complete.obs", method = c("pearson"))
test
```


```{r}
# cannot lump all continuous variables to display into a correlation matrix because of missing NA values... (cut out avg_temp and temp_sd)
con_var <- my_data %>% select(shell_length:heart_rate) %>% drop_na
mydata.cor = cor(con_var, method = c("pearson"))
corrplot(mydata.cor, type = "upper", order = "hclust", tl.col = "black")
palette = colorRampPalette(c("blue", "white", "red")) (20)
heatmap(x = mydata.cor, col = palette, symm = TRUE)
```

```{r}
mydata.rcorr = rcorr(as.matrix(con_var))
mydata.rcorr
corrplot(mydata.rcorr$r, type="upper", order="hclust", tl.col = "black", 
         p.mat = mydata.rcorr$P, sig.level = 0.01, insig = "blank")
```

```{r}
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```


```{r}
mydata.rcorr = rcorr(as.matrix(con_var))
flattenCorrMatrix(mydata.rcorr$r, mydata.rcorr$P)
```

### Individual Pearson Correlation Tests
```{r}
cor.test(my_data$average_weight, my_data$shell_length, method = "pearson")
```
Conclude body weight is positively correlated with shell length

```{r}
#Model_1 <- lmer(trait1 ~ factor1, dataset)
#anova(Model_1, ddf="lme4")
```

### ANOVA

Body weight, shell length, sex and experimental day are not statisically significantly affected.

```{r}
weight_sex.aov <- aov(average_weight ~ sex, data = my_data)
summary(weight_sex.aov)
```
```{r}
weight_day.aov <- aov(average_weight ~ experimental_day, data = my_data)
summary(weight_day.aov)
```
```{r shell length vs sex}
length_sex.aov <- aov(shell_length ~ sex, data = my_data)
summary(length_sex.aov)
```
```{r}
length_day.aov <- aov(shell_length ~ experimental_day, data = my_data)
summary(length_day.aov)
```

## Covariables with Dependent Variables

### Correlation Tests
Please refer to the correlation matrix above. Below are individual tests for body weight and dependent variables.
```{r}
#cor(x, method = "pearson", use = "complete.obs")
#weight <- mydata[, c(7,8,12,13,14)] %>% drop_na("osmolality")
```

```{r}
cor.test(my_data$average_weight, my_data$weight_loss, method = "pearson")
```
Average weight is correlated with weight loss!!!

```{r}
cor.test(my_data$average_weight, my_data$osmolality, method = "pearson")
```

```{r}
cor.test(my_data$average_weight, my_data$heart_rate, method = "pearson")
```

```{r}
cor.test(my_data$average_weight, my_data$average_temp, method = "pearson")
```

```{r}
cor.test(my_data$average_weight, my_data$temp_sd, method = "pearson")
```

### MANOVA

mshapiro.test( )[in the mvnormtest package] can be used to perform the Shapiro-Wilk test for multivariate normality
```{r}
res1_sex.man <- manova(cbind(my_data$weight_loss_percent, my_data$osmolality, my_data$heart_rate) ~ sex, data = my_data)
summary(res1_sex.man)
summary.aov(res1_sex.man)
```
```{r}
res2_sex.man <- manova(cbind(my_data$average_temp, my_data$temp_sd) ~ sex, data = my_data)
summary(res2_sex.man)
summary.aov(res2_sex.man)
```

Experimental day seems to have an effect on dependent variables...

```{r}
res1_day.man <- manova(cbind(my_data$weight_loss_percent, my_data$osmolality, my_data$heart_rate) ~ experimental_day, data = my_data)
summary(res1_day.man)
summary.aov(res1_day.man)
```

```{r}
res2_day.man <- manova(cbind(my_data$average_temp, my_data$temp_sd) ~ experimental_day, data = my_data)
summary(res2_day.man)
summary.aov(res2_day.man)
```

## Independent with Dependent

Questions:
1. Split tests based on the independent variables (3 factors) or dependent variables (5 variables)?
2. How to best time point within heating rate?
3. Do I need to do residual plots?

```{r}
m_weight <- aov(my_data$weight_loss_percent ~ my_data$heating_rate * my_data$tidal_height * my_data$time_point, data = my_data)
summary(m_weight)
```
Need to further process to eliminate 3-way and 2-way interactions since they are insiginificant?

```{r}
m_osmolality <- aov(my_data$osmolality ~ my_data$heating_rate * my_data$tidal_height * my_data$time_point, data = my_data)
summary(m_osmolality)
```

```{r}
m_hr <- aov(my_data$heart_rate ~ my_data$heating_rate * my_data$tidal_height * my_data$time_point, data = my_data)
summary(m_hr)
```

```{r}
m_atemp <- aov(my_data$average_temp ~ my_data$heating_rate * my_data$tidal_height * my_data$time_point, data = my_data)
summary(m_atemp)
```

```{r}
m_tempsd <- aov(my_data$temp_sd ~ my_data$heating_rate * my_data$tidal_height * my_data$time_point, data = my_data)
summary(m_tempsd)
```
## Checking Assumptions

### Normality
Reference: https://cran.r-project.org/web/packages/olsrr/vignettes/residual_diagnostics.html
```{r}
# Weight loss percent
loss_model <- lm(my_data$weight_loss_percent ~ my_data$heating_rate * my_data$tidal_height * my_data$time_point, data = my_data)
ols_plot_resid_qq(loss_model)

# Osmolality
osmolality_model <- lm(my_data$osmolality ~ my_data$heating_rate * my_data$tidal_height * my_data$time_point, data = my_data)
ols_plot_resid_qq(osmolality_model)

# Heart rate
hr_model <- lm(my_data$heart_rate ~ my_data$heating_rate * my_data$tidal_height * my_data$time_point, data = my_data)
ols_plot_resid_qq(hr_model)

# Average temp
temp_model <- lm(my_data$average_temp ~ my_data$heating_rate * my_data$tidal_height * my_data$time_point, data = my_data)
ols_plot_resid_qq(temp_model)
```

```{r}
ols_test_normality(loss_model)
ols_test_normality(osmolality_model)
ols_test_normality(hr_model)
ols_test_normality(temp_model)
```

```{r}
Correlation between observed residuals and expected residuals under normality.

model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_correlation(model)
```


### Homogeneity of Variances

```{r}
leveneTest(my_data$weight_loss_percent ~ interaction(my_data$heating_rate, my_data$tidal_height, my_data$time_point), data = my_data)


```

### Post-hoc Tests

```{r}

```


## References
+ correlation matrix: http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software
+ three-way ANOVA: https://www.r-bloggers.com/2017/02/raccoon-ch-2-4-3-way-anova/